---
title: "Ex4"
author: "Chathura J Gunasekara"
date: "Monday, September 29, 2014"
output: pdf_document
---
1.a
Classify music into six catagories
12,495 music samples 191 predictors.


In this situation, since the sample size is large it is possible to 
set aside a testing data set and training data set.
regardless of the artist, the underlying connection between songs, it is possible to
catagorise the songs to different genres.
since the data shows no large disproposion between the classes ,we can use simple random sample
The response categories were not balanced
(Fig. 1.1), with the smallest segment coming from the heavy metal category
(7%) and the largest coming from the classical category (28%).
do stratified random sampling. the random sampling occurs within each class and should preserve the overall class distribution of the data. 

```{r,echo=TRUE}
music<- read.csv("genresTrain.csv")
attach(music)
#b)
library(AppliedPredictiveModeling)
#summary(music)

barplot(table(music[192]))

#stratified sampling
library(caret)
#install.packages("proxy")
library(proxy)
set.seed(1);trainingRows <- createDataPartition(music$GENRE,p=0.80,list=FALSE)
head(trainingRows)
train_music<-music[trainingRows,]
train_classes<-GENRE[trainingRows]
test_music <-music[-trainingRows,]
test_classes <- GENRE[-trainingRows]



```
Using maximum dissimilarity sampling in caret package. the data will be split on the basis of the predictor values.
```{r, echo=TRUE}

## A random sample of 5 data points
startSet <- sample(1:dim(train_music)[1], 5)
samplePool <- train_music[-startSet, ]
start <- train_music[startSet, ]
newSamp <- maxDissim(start, samplePool, n = 4)

```
4.2.a
```{r,echo=TRUE}
library(AppliedPredictiveModeling)
data(permeability)
per_data <-cbind(permeability,fingerprints)

```
This data set is small. so do do bootstrapping
k fold cross validation
repeated cross validation 
is used.

```{r,echo=TRUE}
library(caret)
set.seed(1)
repeatSplits <-createDataPartition(permeability,p=0.8,times=3)
str(repeatSplits)
set.seed(1)
cvSplits <-createFolds(permeability,k=10,returnTrain=TRUE)

str(cvSplits)


```


4.3.
This data set contains information about a chemical manufacturing process,
in which the goal is to understand the relationship between the process and
the resulting final product yield. Raw material in this process is put through
a sequence of 27 steps to make the final pharmaceutical product. The starting
material is generated from a biological unit and has a range of quality
and characteristics. The objective in this project was to develop a model to
predict percent yield of the manufacturing process. The data set consisted of
177 samples of biological material for which 57 characteristics were measured.

(a)A parsimonious model is a model that accomplishes a desired level of explanation or prediction with as few predictor variables as possible.
THe "one standard error method"



```{r,echo=TRUE}
library(AppliedPredictiveModeling)
data(ChemicalManufacturingProcess)
Rsqared <- cbind((1:10),c(0.444,0.5,0.533,0.545,0.542,0.537,0.534,0.534,0.520,0.507))
plot(Rsqared,xlab="Components",ylab="Resampled R-Squared")







```
Models with components 1 to 4 increase R squared. then after 4 it decreases becasue over fitting.
Numercically optimal value is 0.545 its SD is 0.0308
onestandar_error = 0.545-0.0308
0
.533 > 0.5142 which is within one standard deviation
so 3 Components is enough to model.

b)
mean <- c(0.444,0.5,0.533,0.545,0.542,0.537,0.534,0.534,0.520,0.507)
tol <- (mean-0.545)/0.545
tol*100

54.5-10
44.5
2 is the optimal number of PLS components.

c)
knn
R Squared is between range and prediction time is less than SVM and Random forest. No much defferece in time when comaparing with other models.
d)
start with SVM,boosted trees
investigate simpler models that are less opaqure 
consider using the simlest model that reasonably approximates the performance of more complex methods.

4.4
```{r,echo=TRUE}
library(caret)
data(oil)
str(oilType)
table(oilType)
```
(a)
```{r,echo=TRUE}
sample(fattyAcids,7,replace=TRUE,prob=NULL )






