---
title: "Ex12"
author: "Chathrua Gunasekara"
date: "November 29, 2014"
output:
  word_document:
    fig_width: 4.5
---


Q3. 1. Data Exploration
```{r,echo=FALSE}
rm(list=ls())
library(C50)
library(caret)
data(churn)
dataset <-rbind(churnTrain,churnTest)
X <- dataset[,-20]
```
Remove Near Zero Variance predictors

```{r, echo=FALSE}
nzro <- nearZeroVar(X)
nzro
X <-X[,-6]
```
Removed predictor V6 because the near zero variance.


```{r, echo=FALSE}
numerics <- X[,-c(1,3,4,5)]
corrrelation <- cor(numerics)
library(corrplot)
corrplot(corrrelation, order = "hclust")
numerics <- numerics[,-findCorrelation(cor(numerics))]

prProc <- preProcess(numerics,method=c("center","scale"))
X <- predict(prProc, numerics)
Y <-dataset[,20]
```
Removed the catagorical predictors and high correlated numerical predictors.
Then the data set was pre processed with "centering" and "scaling"

```{r,echo=FALSE}
barplot(table(Y))
trainIndex <- createDataPartition(Y, p = .75,list = FALSE,times = 1)
trainX <-X[trainIndex,]
trainY <-Y[trainIndex]
testX <- X[-trainIndex,]
testY <- Y[-trainIndex]

```
The data set was splited to 75% training set and rest to testing set with stratified sampling beucase the class imbalance in the data set.

2.Fitting models and evaluating the model using ROC. Kappa statistic is used because th class imbalance.
```{r,echo=FALSE}
ctrl <- trainControl(method = "LGOCV",
                     summaryFunction = twoClassSummary,
                     classProbs = TRUE,
                     savePredictions = TRUE)
```

i. Logistic Regression classification model
```{r,echo=FALSE}

glmModel <- train(trainX,y = trainY,method = "glm",metric = "ROC",trControl = ctrl)
glmModel

glmPred <- predict(glmModel,testX)
confusionMatrix(data=glmPred,reference = testY)
Roc <- roc(response = glmModel$pred$obs,
                                       predictor = glmModel$pred$yes,
                                       levels = (levels(glmModel$pred$obs)))
#plot(Roc, legacy.axes = TRUE)
auc(Roc)
predLda<-predict(glmModel,testX)
confusionMatrix(data=predLda,reference = testY)

```
ii.Linear Discriminate Analysis
```{r,echo=FALSE}
ldaModel<- train(x = trainX,y = trainY,method = "lda",metric = "ROC",trControl = ctrl)
ldaModel
Roc <- roc(response = ldaModel$pred$obs,
           predictor = ldaModel$pred$yes,
           levels = (levels(ldaModel$pred$obs)))
#plot(Roc, legacy.axes = TRUE)
auc(Roc)

predLda<-predict(ldaModel,testX)
confusionMatrix(data=predLda,reference = testY)

```
iii. Partial Least Squrare Discriminate Analysis
```{r,echo=FALSE}

plsdaModel <- train(x = trainX,y = trainY,
                                     method = "pls",
                                     tuneGrid = expand.grid(.ncomp = 1:2),
                                     #preProc = c("center","scale"),
                                     metric = "ROC",
                                     trControl = ctrl)
plsdaModel


Roc <- roc(response = plsdaModel$pred$obs,
           predictor = plsdaModel$pred$yes,
           levels = (levels(plsdaModel$pred$obs)))
#plot(Roc, legacy.axes = TRUE)
auc(Roc)

predLda<-predict(plsdaModel,testX)
confusionMatrix(data=predLda,reference = testY)

plsImpGrant <- varImp(plsdaModel, scale = FALSE)
plsImpGrant
```
iv. Penalized Models
```{r,echo=FALSE}
glmnGrid <- expand.grid(.alpha = c(0, .1 ),.lambda=seq(0.1,0.2,length=5))
lmnTuned <- train(trainX,y = trainY,method = "glmnet",  tuneGrid = glmnGrid,metric = "ROC",trControl = ctrl)
lmnTuned

predLda<-predict(lmnTuned,testX)
confusionMatrix(data=predLda,reference = testY)
#Roc <- roc(response = lmnTuned$pred$obs,
 #          predictor = lmnTuned$pred$yes,
  #         levels = (levels(lmnTuned$pred$obs)))
#plot(Roc, legacy.axes = TRUE)
#auc(Roc)
```


v. Nearest Shrunken centroids Model
```{r,echo=FALSE}
library(pamr)
nscGrid <- data.frame(.threshold = 0:25)
nscTuned <- train(x = trainX,
                                       y = trainY,
                                       method = "pam",
                                       #preProc = c("center", "scale"),
                                       tuneGrid = nscGrid,
                                       metric = "ROC",
                                       trControl = ctrl)

nscTuned
predLda<-predict(nscTuned,testX)
confusionMatrix(data=predLda,reference = testY)


predictors(nscTuned)
varImp(nscTuned, scale = FALSE)

```



